---
title: "Untitled"
author: "Chris Peralta"
date: "May 10, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(max.print = 10)
library(tidyverse)
library(janitor)
library(lubridate)
library(naniar)
library(labelled)
library(stringr)

source('functions/word_count.R')
```


For cleaning this data, I'm going to take some preliminary steps and then drop the variables


```{r cars}
tast_db <- haven::read_sav("Data/tastdb-exp-2016.sav")


tast_db <- tast_db %>% 
  convert_to_NA(c('')) %>% #some missing values are coded with empty strings
  remove_empty(c('rows', "cols"))#removes adult4, no data in this col
```


Use tonmod as it is the imputed variable that standardizes all of the tonnage measurements. One of the database creators, says that it is the least reliable imputed variable but I will use it rather than attempt to reimpute the variable more accurately.
http://www.slavevoyages.org/voyage/understanding-db/methodology-19

-Removed the data sources
-Removed evgreen (another data source)
-For voyage itinerary use imputed variables

-Removed intended
- 

```{r}
glimpse(tast_db)

tast <- tast_db %>% 
  select(-starts_with('source')) %>% 
  select(-tontype, -tonnage, -evgreen) %>% 
  select(-starts_with('slin')) %>% 
  set_na_values(fate4 = 4) %>% 
  user_na_to_na() %>% 
  mutate(fate4 = to_factor(fate4))



prop_miss(tast) # 77.6% of values are missing

var_missing_under_50 <- tast %>%
  miss_var_summary() %>% 
  arrange(desc(pct_miss)) %>% 
  filter(pct_miss < 53.6) %>% 
  pull(variable) 
  
tast1 <- tast %>% 
  select(one_of(var_missing_under_50), -national)
# tast <- tast %>% 
#   select(-one_of(setdiff(var_missing_under_50, c('Date_land1', 'dateland1', 
#                                                    'Date_land2',
#                                      'dateland2', 'Date_land3', 'dateland3', 
#                                      'dateend','Date_end', 'dateleftafr','Date_leftafr',
#                                      'Date_buy', 'datebuy', 'Date_depam', 'datedepam'))),
#          -national) # Use natinimp, the imputed version
#                                             #keeping these dates for now

```


Variables of note: shipname, national, tonmod (Tonnage standardized on Briti sh measured tons,1773-1835), yrcons, guns, placcons, owner_

note: xmimpflag is a variable grouping the voyages for the purpose of imputing slave counts

tonmod was dropped, above,  with 53.6% of it's values missing, consider adding some of the variables back. 



Here, I will set the columns to the proper data structures
```{r}
tast1 %>% glimpse()
#factors: captaina, ownera, shipname
#dates: datedep, dateland1
#to_factor: deptregimp, deptregimp1, starts_with("m"), natinimp, plac1tra, portdep, ptdepimp, regdis1, regem1, rig, sla1port

temp <- tast1 %>% 
  select(-starts_with('fate')) %>%
  mutate_if(is.labelled, to_factor) %>% 
  mutate_at(vars(captaina, ownera, shipname), factor) %>% 
  mutate_if(is.character, ymd) %>% 
  select(-`filter_$`)

tast1 <- tast1 %>% 
  select(starts_with('fate')) %>% 
  bind_cols(temp)


tast1 %>% select(starts_with("datedep")) %>% miss_var_summary() 
tast1 %>% select(starts_with("datedep")) %>% View() 
#datedep is constructed from datedepa,b,c

tast1 %>% select(starts_with("datarr")) %>% miss_var_summary() 
tast1 %>% select(starts_with("datarr"), dateland1) %>% View() 
#dateland1 is constructed from datarr32, 33, 34

tast1 %>% 
  mutate_at(.vars = vars(matches("^year\\d{1,2}")), .funs = funs(to_character)) %>% 
  select(starts_with('year')) 

```



Need to look at years
```{r pressure, echo=FALSE}
tast1 %>% glimpse()



tast %>% select(matches('^year\d')) %>% glimpse() #

tast1 %>% 
  count(factor(fate4)) %>% 
  ggplot() + geom_col(aes(x = fct_reorder(`factor(fate4)`, -n), y = n))


tast1 %>% miss_var_summary() %>% arrange(desc(n_miss)) %>% View()

tast1 %>% 
  ggplot(aes(x = yeardep)) + 
  geom_bar()


```



Analyzing the 'fate' variable for possible, more generalized, recodings

fate: 
   Condemned- By the general practice of the law of nations, a sentence of condemnation is, at present, generally deemed necessary in order to divest the title of a vessel taken as a prize. Until this has been done the original owner may regain his property, although the ship may have been in possession of the enemy twenty-four hours, or carried infra praesidia. https://legal-dictionary.thefreedictionary.com/condemnation


See regroupings.txt for tast$fate recodings

```{r}
tast1 %>% 
  mutate(fate_char = to_character(fate)) %>% 
  word_count(fate_char)

#keywords: completed, sold, captured, destroyed shipwrecked

```

Some EDA. Below, I split the 'fate' variable into 'As intended' and 'Not as intended'. I'm currently trying to find out which variables are likely to contribute to a successful transatlantic voyage, to better understand the data. At fate == 1, the value is 'Voyage completed as intended, which is why I've chosen to recode the variable in this manner. 
The '_re' suffix is shorthand for 'recoded'. 
```{r}
tast_re <- tast1 %>% 
  mutate(fate_re = case_when(fate == 1 ~ 'As intended', 
                             fate >= 1 ~ 'Not as intended')) %>% 
  mutate(fate_re = as.factor(fate_re))


qplot(x = yeardep, 
      data = tast_re,
      geom = 'bar')

tast_re %>% 
  mutate(row_missing = prop_miss_row(tast1)) %>% 
  group_by(yeardep) %>% 
  summarize(year_missing = mean(row_missing)) %>% 
  qplot(x = yeardep, y = year_missing, 
        data = ., 
        geom = c('point', 'smooth')) + 
  geom_vline(xintercept = 1807) + 
  geom_vline(xintercept =  1836) + 
  annotate('text', label = 'Great Britain outlawed slavery 1807', 
           x = 1800, y = .45, angle = 90) + 
  annotate('text', label = 'Portugal outlawed slavery 1836', 
           x = 1829, y = .45, angle = 90)
  # The data is not missing at random. Missing data varies by year

tast_re %>% 
  mutate(row_missing = prop_miss_row(tast1)) %>% 
  group_by(datedepb) %>% 
  summarize(month_missing = mean(row_missing)) %>% 
  filter(!is.na(datedepb)) %>% 
  qplot(x = factor(datedepb), y = month_missing,  
        data = ., 
        geom = c('point')) +
  scale_y_continuous(limits = c(0,.4))
  # Looks like missing data barely increases from june to december, which is during
  # atlantic hurricane season. This may indicate that poor weather could be affecting the 
  # outcome of the journey, thus making it harder to collect data 

  
  
tast_re %>% 
  mutate(row_missing = prop_miss_row(tast_re)) %>% 
  group_by(yeardep, fate_re) %>% 
  summarize(year_missing = mean(row_missing)) %>% 
  qplot(x = yeardep, y = year_missing, color = fate_re,
        data = ., 
        geom = c('point', 'smooth')) + 
  geom_vline(xintercept = 1807) + 
  geom_vline(xintercept =  1836) + 
  annotate('text', label = 'Great Britain outlawed slavery 1807', 
           x = 1800, y = .45, angle = 90) + 
  annotate('text', label = 'Portugal outlawed slavery 1836', 
           x = 1829, y = .45, angle = 90) 
  # Obs with 'Voyage completed as intended', are missing less data





qplot(x = yeardep, fill = fate_re,
      data = tast_re, 
      geom = 'histogram') + 
  scale_fill_manual(values = c('Blue', 'Red'))
  # appears to be trend between year departed and fate, likely because of the partially known     fates in 'Not as intended'

tast_re %>% 
  filter(yeardep < 1600) %>% 
  select(yeardep, fate) %>% 
  mutate(fate = to_character(fate)) %>% 
  count(fate) %>% 
  arrange(desc(n)) %>% 
  View
  # 'Not as intended' dominates the previous graph, this appears to be because of                 partially-known fates 


qplot(x = slaximp, bins = 40,
      data = tast_re, 
      geom = 'histogram')
qplot(x = slamimp, bins = 40, 
      data = tast_re, 
      geom = 'histogram')
qplot(x = slaximp, y = slamimp, alpha = .1, 
      data = tast_re, 
      geom = 'point')
  # slaximp: slaves embarked, slamimp: slaves disembarked

tast_re %>% 
  filter(tonmod < 600, slaximp < 1000) %>% 
  qplot(x = tonmod, y = slaximp, alpha = .00000000001,
        data = ., 
        geom = c('jitter', 'smooth'))
  # It appears that there is some relationship between tonnage and number of slaves embarked,     it also appears that there is a discernible pattern in the imputation

tast_re %>% 
  mutate(row_missing = prop_miss_row(tast_re)) %>% 
  group_by(yeardep, fate4) %>% 
  summarize(year_missing = mean(row_missing)) %>% 
  qplot(x = yeardep, y = year_missing, facets = ~fate4,
        data = ., 
        geom = c('point', 'smooth')) + 
  geom_vline(xintercept = 1807) + 
  geom_vline(xintercept =  1836) + 
  annotate('text', label = 'Great Britain outlawed slavery 1807', 
           x = 1800, y = .45, angle = 90) + 
  annotate('text', label = 'Portugal outlawed slavery 1836', 
           x = 1829, y = .45, angle = 90) 


```


As you can see below, the data is not missing at random. There is more missing data for earlier years and less for later years, until 1807 when Great Britain made it illegal to engage in the slave trade. The same year, the US passed The Act Prohibiting Importation of Slaves of 1807
```{r}
# The data is not missing at random.
tast1 %>% 
  mutate(row_missing = prop_miss_row(tast1)) %>% 
  group_by(yeardep) %>% 
  summarize(year_missing = mean(row_missing)) %>% 
  qplot(x = yeardep, y = year_missing, 
        data = ., 
        geom = c('smooth', 'point')) + 
  geom_vline(xintercept = 1807) + 
  geom_vline(xintercept =  1836) + 
  annotate('text', label = 'Great Britain outlawed slavery 1807', 
           x = 1800, y = .45, angle = 90) + 
  annotate('text', label = 'Portugal outlawed slavery 1836', 
           x = 1829, y = .45, angle = 90)


```


Analyze trip length / duration, 

```{r}
tast1 %>% glimpse()


```





Variables most likely to predict fate: 
yeardep, datedep, datedepb, tonmod, dateland, voy1imp, voy2imp, port began, port endend, slaximp, natinimp, shipname, season

ptdepimp, mjslptimp can't be used as factors, because there will be values in the training set that aren't present in the testing set, need to convert the strings to numeric. Also consider calculating the distance between ports

playing around with modelling

```{r}
library(caret)
train <- createDataPartition(tast_re$fate_re, p = .6, list = FALSE)
training <- tast_re[train, ]
testing <- tast_re[-train, ]

mod2 <- glm(fate_re ~ yeardep + tonmod + natinimp + slamimp + slaximp + 
                      slaximp*slamimp, 
            data = training, family = binomial(link = 'logit'), na.action = na.omit)

mod1 <- polr(fate2 ~ dateland1, 
            data = training, na.action = na.omit, Hess = TRUE) 



testing %>% 
  mutate(pred = predict(mod2, ., type = 'response')) %>% 
  ggplot(aes(x = pred, fill = fate_re)) + 
  geom_histogram()


p <- predict(mod2, testing, type="response")

p_class <- if_else(p > 0.7729328, 'Success', 'Not Success')

confusionMatrix(as.factor(p_class), testing$fate_re)


#Caret syntax
mod <- train(fate_re ~ dateland1, 
             data = training, na.action = na.omit, 
             method = 'glm', family = 'binomial')


```

