---
title: "Untitled"
author: "Chris Peralta"
date: "May 10, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(max.print = 10)
library(tidyverse)
library(janitor)
library(lubridate)
library(naniar)
library(labelled)
library(stringr)

source('functions/word_count.R')
```


For cleaning this data, I'm going to take some preliminary steps and then drop the variables


```{r cars}
tast_db <- haven::read_sav("Data/tastdb-exp-2016.sav")


tast_db <- tast_db %>% 
  convert_to_NA(c('')) %>% #some missing values are coded with empty strings
  remove_empty(c('rows', "cols"))#removes adult4, no data in this col
```


Use tonmod as it is the imputed variable that standardizes all of the tonnage measurements. One of the database creators, says that it is the least reliable imputed variable but I will use it rather than attempt to reimpute the variable more accurately.
http://www.slavevoyages.org/voyage/understanding-db/methodology-19

-Removed the data sources
-Removed evgreen (another data source)
-For voyage itinerary use imputed variables

-Removed intended
- 

```{r}
glimpse(tast_db)

tast <- tast_db %>% 
  select(-starts_with('source')) %>% 
  select(-tontype, -tonnage, -evgreen) %>% 
  select(-starts_with('slin')) %>% 
  set_na_values(fate4 = 4) %>% 
  user_na_to_na()



prop_miss(tast) # 77.6% of values are missing

var_miss_under_50 <- tast %>%
  miss_var_summary() %>% 
  arrange(desc(pct_miss)) %>% 
  filter(pct_miss < 50) %>% 
  pull(variable)
  
tast1 <- tast %>% 
  select(one_of(var_miss_under_50), -national)
# tast <- tast %>% 
#   select(-one_of(setdiff(var_miss_over_50, c('Date_land1', 'dateland1', 
#                                                    'Date_land2',
#                                      'dateland2', 'Date_land3', 'dateland3', 
#                                      'dateend','Date_end', 'dateleftafr','Date_leftafr',
#                                      'Date_buy', 'datebuy', 'Date_depam', 'datedepam'))),
#          -national) # Use natinimp, the imputed version
#                                             #keeping these dates for now

```


Variables of note: shipname, national, tonmod (Tonnage standardized on Briti sh measured tons,1773-1835), yrcons, guns, placcons, owner_

note: xmimpflag is a variable grouping the voyages for the purpose of imputing slave counts

fate: 
   Condemned- By the general practice of the law of nations, a sentence of condemnation is, at present, generally deemed necessary in order to divest the title of a vessel taken as a prize. Until this has been done the original owner may regain his property, although the ship may have been in possession of the enemy twenty-four hours, or carried infra praesidia. https://legal-dictionary.thefreedictionary.com/condemnation
   1- voyage completed as intended
   40- sold
   49- Sold slaves in Americas - subsequent fate unknown 
   68- Sold in the Americas after disembarking slaves
   70- Sold in the Americas - not known whether ship brought slaves 
   77- Arrived in Africa, subsequent fate unknown 
   86- Ship returned direct to home port after selling slaves to another vessel 
   88- Sold prematurely in Europe after disembarking slaves in the Americas 
   98- Ship and slaves sold in Africa
   188- Temporarily detained by the Dutch: Voyage allowed to continue with slaves. 

tonmod was dropped with 53.6% of it's values missing, consider adding some of the variables back. 



Here, I will set the columns to the proper data structures
```{r}
tast1 %>% glimpse()
#factors: captaina, ownera, shipname
#dates: datedep, dateland1
#to_factor: deptregimp, deptregimp1, starts_with("m"), natinimp, plac1tra, portdep, ptdepimp, regdis1, regem1, rig, sla1port

temp <- tast1 %>% 
  select(-starts_with('fate')) %>%
  mutate_if(is.labelled, to_factor) %>% 
  mutate_at(vars(captaina, ownera, shipname), factor) %>% 
  mutate_if(is.character, ymd) %>% 
  select(-`filter_$`)

tast1 <- tast1 %>% 
  select(starts_with('fate')) %>% 
  bind_cols(temp)


tast1 %>% select(starts_with("datedep")) %>% miss_var_summary() 
tast1 %>% select(starts_with("datedep")) %>% View() 
#datedep is constructed from datedepa,b,c

tast1 %>% select(starts_with("datarr")) %>% miss_var_summary() 
tast1 %>% select(starts_with("datarr"), dateland1) %>% View() 
#dateland1 is constructed from datarr32, 33, 34

tast1 %>% 
  mutate_at(.vars = vars(matches("^year\\d{1,2}")), .funs = funs(to_character)) %>% 
  select(starts_with('year')) 

```



Need to look at years
```{r pressure, echo=FALSE}
tast1 %>% glimpse()



tast %>% select(matches('^year\d')) %>% glimpse() #

tast1 %>% 
  count(factor(fate4)) %>% 
  ggplot() + geom_col(aes(x = fct_reorder(`factor(fate4)`, -n), y = n))


tast1 %>% miss_var_summary() %>% arrange(desc(n_miss)) %>% View()

tast1 %>% 
  ggplot(aes(x = yeardep)) + 
  geom_bar()


```



Analyzing the 'fate' variable for possible, more generalized, recodings
```{r}
tast1 %>% 
  mutate(fate_char = to_character(fate)) %>% 
  possibly_word_count(fate_char) %>% View()
#completed, sold, 
  
```

Some EDA. Below, I split the 'fate' variable into 'As intended' and 'Not as intended'. I'm currently trying to find out which variables are likely to contribute to a successful transatlantic voyage, to better understand the data. At fate == 1, the value is 'Voyage completed as intended, which is why I've chosen to recode the variable in this manner. 
The '_re' suffix is shorthand for 'recoded'. 
```{r}
tast_re <- tast1 %>% 
  mutate(fate_re = case_when(fate == 1 ~ 'As intended', 
                             fate >= 1 ~ 'Not as intended')) %>% 
  mutate(fate_re = as.factor(fate_re))


qplot(x = yeardep, fill = fate_re, data = tast_re, geom = 'histogram')
  #appears to be trend between year departed and fate

# The data is not missing at random.
tast1 %>% 
  mutate(row_missing = prop_miss_row(tast1)) %>% 
  group_by(yeardep) %>% 
  summarize(year_missing = mean(row_missing)) %>% 
  qplot(x = yeardep, y = year_missing, 
        data = ., 
        geom = 'point') + 
  geom_vline(xintercept = 1807) + 
  geom_vline(xintercept =  1836) + 
  annotate('text', label = 'Great Britain outlawed slavery 1807', 
           x = 1800, y = .45, angle = 90) + 
  annotate('text', label = 'Portugal outlawed slavery 1836', 
           x = 1829, y = .45, angle = 90)


```


As you can see below, the data is not missing at random. There is more missing data for earlier years and less for later years, until 1807 when Great Britain made it illegal to engage in the slave trade. The same year, the US passed The Act Prohibiting Importation of Slaves of 1807
```{r}
# The data is not missing at random.
tast1 %>% 
  mutate(row_missing = prop_miss_row(tast1)) %>% 
  group_by(yeardep) %>% 
  summarize(year_missing = mean(row_missing)) %>% 
  qplot(x = yeardep, y = year_missing, 
        data = ., 
        geom = c('smooth', 'point')) + 
  geom_vline(xintercept = 1807) + 
  geom_vline(xintercept =  1836) + 
  annotate('text', label = 'Great Britain outlawed slavery 1807', 
           x = 1800, y = .45, angle = 90) + 
  annotate('text', label = 'Portugal outlawed slavery 1836', 
           x = 1829, y = .45, angle = 90)


```




playing around with modelling

```{r}
train <- createDataPartition(tast_re$fate_re, p = .6, list = FALSE)
training <- tast_re[train, ]
testing <- tast_re[-train, ]

mod2 <- glm(fate_re ~ dateland1, data = training, family = binomial(link = 'logit'), na.action = na.omit)

mod1 <- polr(fate2 ~ dateland1, 
            data = training, na.action = na.omit, Hess = TRUE) 



testing %>% 
  mutate(pred = predict(mod2, ., type = 'response')) %>% 
  ggplot(aes(x = pred, fill = fate_re)) + 
  geom_histogram()


p <- predict(mod2, testing, type="response")

p_class <- if_else(p > 0.7729328, 'Success', 'Not Success')

confusionMatrix(as.factor(p_class), testing$fate_re)


#Caret syntax
mod <- train(fate_re ~ dateland1, 
             data = training, na.action = na.omit, 
             method = 'glm', family = 'binomial')


```

